Behavior Driven Development in Haskell

### 2012-05-28 paper.tex
Remove trailing whitespace
Introduce implicitly quantified properties
Describe how properties are tested

\noindent Evidence in comments are automatically
tested by the Doctest command. We also provided a library
to do this job. Its implementation will discussed in Section~\ref{sec:doctest}.

@@

To test properties, we use GHCi.  Each property is enclosed within
``\verb|quickCheck (|'' and ``\verb|)|'', and GHCi is used to evaluate
that.  GHCi output is used to decide, whether the property succeeded,
or failed.

If a property contains free variables, we need to add lambda
abstraction prefix before prepending the `quickCheck'' string.  To
check if properties contain free variables, we also use GHCi for easy
implementation at this moment.  GHCi reports free variables if found:

\begin{quote}
\small
\begin{verbatim}
Prelude Data.List> sort xs == (sort . sort) (xs :: [Int]))
<interactive>:1:6: Not in scope: `xs'
<interactive>:1:24: Not in scope: `xs'
\end{verbatim}
\end{quote}

\noindent In this case, we should prepend the ``\verb|\xs -> |" string
to the property.

\begin{quote}
    TBD: describe the algorithm

    Below is the schematic algorithm, from my original mail to kazu.
    Can we somehow get rid of haskell-src, and hint, and only describe
    the algorithm?
\end{quote}

\noindent \verb|xs| is a free variable here.  So we would need to
close that term by transforming it to:

@@

\noindent I imagine this could work by performing those steps:

\begin{enumerate}

    \item
        parse the property with e.g. haskell-src (never used it, so
        I'm not sure if it allows you to parse code snippets)

    \item
        Load the module containing the DocProperty with hint, and try
        to look up all variables used by the property (\verb|sort| and
        \verb|xs| for the example above).

    \item
        Close the term over all variables, for which lookup failed,
        load it with hint and test it with quickCheck.
\end{enumerate}

>>>>

\noindent We call such properties \emph{implicitly quantified}.  The
given property is implicitly quantified over \verb|xs|.

Evidence in comments are automatically
tested by the Doctest command. We also provided a library
to do this job. Its implementation will discussed in Section~\ref{sec:doctest}.

@@

\begin{quote}
    sol: Should we use verify instead of test here?
\end{quote}

\noindent QuickCheck and GHCi are used to test properties.  Before we
test a property, we bring the module that contains the Haddock comment
from which we extracted the property into scope.

Implicitly quantified properties need to be closed before they can be
tested.  This requires the following steps.
\begin{enumerate}
    \item
        Find all free variables of the property
    \item
        Try to lookup each free variable in the module context
    \item
        Close the property over all free variables, for which lookup
        failed
\end{enumerate}

\noindent Here is an example:

\begin{quote}\small\begin{verbatim}
sort xs == (sort . sort) (xs :: [Int]))
\end{verbatim}\end{quote}

\noindent Both \verb|xs| and \verb|sort| are free variables, but
\verb|sort| is bound by the module context.  So we need to close the
term over \verb|xs|.

@@
